---
title: "Project Proposal"
author: "Bob Ding, Lynn Fan, Alice Jiang"
date: "11/08/2018"
output: html_document
---

```{r setup, message=F,include=F}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
```

```{r}
library("cowplot")
library("tibble")
library("dplyr")
library("ggplot2")
library("broom")
library("knitr")
```

## Introduction

## Data
```{r}
data <- read.csv("./redwine_quality.csv")
glimpse(data)
```
The data has 12 variables:


###**Response Variable:**
  + **quality**: the quality of the wine
  
###**Explanatory Variables:**
  + **fixed acidity**: the amount of acid in wine that's not volatile
  + **volatile acidity**: the amount of acid in wine that's volatile
  + **citric acid**: the quantity of citric acid, which add freshness flavor to wine
  + **residual sugar**: the amount of sugar left after fermatation stops
  + **chlorides**: the amount of salt in the wine
  + **free sulfure dioxide**: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion
  + **total sulfur dioxide**: amount of free and bound forms of S02
  + **density**: the density, which is related to proportion of water and other solvent.
  + **pH**: the indicator showing the acidicity or basic property. ranging from 0 to 14.
  + **sulphates**: a wine additive which can contribute to sulfur dioxide gas (S02) levels
  + **alcohol**: percent of alcohol


## Analysis
## Reference
## Appendix

<!-- Label each question and question part (if applicable) clearly using R Markdown syntax. -->



