

# Ordinal Logistic Regression
Because the response variable, quality, is ordinal and has a natural ordering for categories 0,1,...,10, we believe ordinal logistic regression might be a good model.

```{r}
# create separate indicator variables for each level of quality to use in binned plots
data <- data %>% mutate(qua0 = if_else(quality_dup=="0",1,0), 
                        qua1 = if_else(quality_dup=="1",1,0),
                        qua2 = if_else(quality_dup=="2",1,0),
                        qua3 = if_else(quality_dup=="3",1,0),
                        qua4 = if_else(quality_dup=="4",1,0),
                        qua5 = if_else(quality_dup=="5",1,0),
                        qua6 = if_else(quality_dup=="6",1,0),
                        qua7 = if_else(quality_dup=="7",1,0),
                        qua8 = if_else(quality_dup=="8",1,0),                            
                        qua9 = if_else(quality_dup=="9",1,0),
                        qua10 = if_else(quality_dup=="10",1,0))
                          
``` 

```{r}
# full model
full.model <- polr(quality_dup ~ volatile.acidityCent + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent, data=data)
summary(full.model)
```

```{r}
# view the model
kable(tidy(full.model),format="markdown",digits=3)
``` 

We see there are 5 possible intercepts/models. We also observe the magnitude of the test statistic for densityCent is significantly higher than the test statistic magnitude of other variables.

```{r}
# parameter estimates with confidence level
confint.default(full.model)
``` 

If the 95% CI does not cross 0, the parameter estimate is statistically significant. The CI for volatile.acidityCent, citric.acid, chloridesCent, sulphatesCent, densityCent, pHCent and alcoholCent do not include 0, so we take them as significant and remove residual.sugarCent and free.sulfur.dioxideCent from the model.


```{r}
# obtain p-values
tidy(coeftest(full.model))
```

The p-values of these variables confirm the results of the confidence interval test that only two variables will be removed from the model: residual.sugarCent and free.sulfur.dioxideCent.


```{r}
# new model
model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + sulphatesCent + pHCent + densityCent + alcoholCent, data=data)
summary(model)
```

```{r}
# confidence interval test of the new model
confint.default(model)
```

densityCent contains zero in its 95% CI. Therefore, we remove variable from the updated model and further proceed with more analysis.

```{r}
# new model without densityCent
final.model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + sulphatesCent + pHCent + alcoholCent, data=data)
summary(final.model)
```
The variable sulphatesCent shows concerning data. Based on our current model, For one unit increase in sulphates, the odds of the wine falling at or below quality i multiply by a factor of 15.401872706, given that all other variables are held constant.The 95% confidence interval is between 7.64 and 31.63. This odds change is pretty dramatic, especially when we compare sulphatesCent's coefficient with other coefficients in this model. Therefore, we need to look at sulphatesCent more closely. 

```{r}
par(mfrow=c(2,3))
arm::binnedplot(x=data$sulphatesCent,y=data$resid3,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid4,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid5,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid6,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid7,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid8,xlab="sulphatesCent")
``` 
There are some slight nonrandom patterns in binned residual plots and the exploratory analysis shows that the data has some outliers. To improve on this model, we try to log transform sulphatesCent and see if anything improves.

```{r}
data <- data %>% mutate(logsulphates = log(sulphates))
final.model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + logsulphates + pHCent + alcoholCent, data=data)
summary(final.model)
```

```{r}
# confidence interval test
confint.default(final.model)
```

We will use this as the final ordianl logistic regression model.

```{r}
# view the final model
kable(tidy(final.model))
```

$log(\frac{P(quality\leq3)}{P(quality>3)})=-7.3076570-[-3.4719465*volatile.acidityCent - 0.9100354*citric.acid -5.6128331*chloridesCent + 2.4960668*logsulphates	- 1.6781013	*pHCent + 0.9433494*alcoholCent]$
$log(\frac{P(quality\leq4)}{P(quality>4)})=-5.3535918-[-3.4719465*volatile.acidityCent - 0.9100354*citric.acid -5.6128331*chloridesCent + 2.4960668*logsulphates	- 1.6781013	*pHCent + 0.9433494*alcoholCent]$
$log(\frac{P(quality\leq5)}{P(quality>5)})=-1.6544846-[-3.4719465*volatile.acidityCent - 0.9100354*citric.acid -5.6128331*chloridesCent + 2.4960668*logsulphates	- 1.6781013	*pHCent + 0.9433494*alcoholCent]$
$log(\frac{P(quality\leq6)}{P(quality>6)})=1.1930150-[-3.4719465*volatile.acidityCent - 0.9100354*citric.acid -5.6128331*chloridesCent + 2.4960668*logsulphates	- 1.6781013	*pHCent + 0.9433494*alcoholCent]$
$log(\frac{P(quality\leq7)}{P(quality>7)})=4.2252323-[-3.4719465*volatile.acidityCent - 0.9100354*citric.acid -5.6128331*chloridesCent + 2.4960668*logsulphates	- 1.6781013	*pHCent + 0.9433494*alcoholCent]$

```{r}
# get CI
ci <- confint(final.model)
```

```{r}
# display odds ratio and CI
exp(cbind(OR = coef(final.model), ci))
```

For one unit increase in volatile.acidity, the odds of the wine falling at or below quality i multiply by a factor of 0.027359812, given that all other variables are held constant.

For one unit increase in citric.acid, the odds of the wine falling at or below quality i multiply by a factor of 0.477221357, given that all other variables are held constant.

For one unit increase in chlorides, the odds of the wine falling at or below quality i multiply by a factor of 0.005199004, given that all other variables are held constant.

For sulphates multiplied by a factor of 2, the odds of the wine falling at or below quality i multiply by a factor of exp(12.134671385log(2)), given that all other variables are held constant.

For one unit increase in pH, the odds of the wine falling at or below quality i multiply by a factor of 0.228196887, given that all other variables are held constant.

For one unit increase in % alcohol content, the odds of the wine falling at or below quality i multiply by a factor of 2.583998874, given that all other variables are held constant.

### Assumptions
```{r}
# calculate predicted probabilities for each observation
predprobs <- data.frame(predict(final.model,type="probs"))
predprobs %>% filter(row_number() <= 3) # print predictions for first three observations
```

```{r}
# predicted categories
data <- data %>% mutate(pred.comp = predict(final.model,type="class")) # get predicted categories
result <- data %>% group_by(pred.comp,quality_dup) %>% summarise(n=n()) # summarise actual vs. predicted
kable(spread(result,quality_dup,n),format="markdown") # print actual vs. predicted categories
```

### Model Diagnostics
```{r}
data <- data %>% mutate(resid3 = qua3-predprobs[,1],
                        resid4 = qua4-predprobs[,2],
                        resid5 = qua5-predprobs[,3],
                        resid6 = qua6-predprobs[,4],
                        resid7 = qua7-predprobs[,5],
                        resid8 = qua8-predprobs[,6])
data$predor <- predict(final.model, data)
```

```{r}
data %>% count(predor,quality_dup)
confusionMatrix(data$predor,data$quality_dup)
```

```{r}
# binned residual plots of volatile.acidityCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid3,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid4,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid5,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid6,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid7,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid8,xlab="volatile.acidityCent")
```

```{r}
# binned residual plots of citric.acid for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$citric.acid,y=data$resid3,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid4,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid5,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid6,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid7,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid8,xlab="citric.acid")
```

```{r}
# binned residual plots of chloridesCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$chloridesCent,y=data$resid3,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid4,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid5,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid6,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid7,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid8,xlab="chloridesCent")
```

The binned residual plots show a possible chloridesCent outlier.

```{r}
# binned residual plots of sulphatesCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$logsulphates,y=data$resid3,xlab="logsulphates")
arm::binnedplot(x=data$logsulphates,y=data$resid4,xlab="logsulphates")
arm::binnedplot(x=data$logsulphates,y=data$resid5,xlab="logsulphates")
arm::binnedplot(x=data$logsulphates,y=data$resid6,xlab="logsulphates")
arm::binnedplot(x=data$logsulphates,y=data$resid7,xlab="logsulphates")
arm::binnedplot(x=data$logsulphates,y=data$resid8,xlab="logsulphates")
```

The binned residual plot shows some potential linear trend and outliers in three categories of quality. After sulphates is log transformed, it still shows nonrandom patterns as the residuals over logsulphate bigger than 0.4 is much larger than that when logsulphates is below 0.4 for high quality wines v. lower quality wines (quality<= and >7)

```{r}
# binned residual plots of pHCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$pHCent,y=data$resid3,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid4,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid5,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid6,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid7,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid8,xlab="pHCent")
```

There is also some linear trends in the binned residual plots for pHCent. We should also tranform this variable.

```{r}
# binned residual plots of alcohol.Cent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$alcoholCent,y=data$resid3,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid4,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid5,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid6,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid7,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid8,xlab="alcoholCent")
```
The binned residual plots of alcoholCent for wines with quality>6 v. those with quality <=6 also shows nonrandom patterns, and we need to take this into consideration.


```{r}
kable(tidy(vif(final.model)))
```

## Additional Work - SulphatesCent
```{r}

# full model
full.model <- polr(quality_dup ~ volatile.acidityCent + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent, data=data)
summary(full.model)
```

```{r}
# view the model
kable(tidy(full.model),format="markdown",digits=3)
``` 

We see there are 5 possible intercepts/models. We also observe the magnitude of the test statistic for densityCent is significantly higher than the test statistic magnitude of other variables.

```{r}
# parameter estimates with confidence level
confint.default(full.model)
``` 

If the 95% CI does not cross 0, the parameter estimate is statistically significant. The CI for volatile.acidityCent, citric.acid, chloridesCent, sulphatesCent, densityCent, pHCent and alcoholCent do not include 0, so we take them as significant and remove residual.sugarCent and free.sulfur.dioxideCent from the model.

```{r}
# new model
model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + sulphatesCent + pHCent + densityCent + alcoholCent, data=data)
summary(model)
```

```{r}
# confidence interval test of the new model
confint.default(model)
```

densityCent contains zero in its 95% CI.

```{r}
# new model without densityCent
final.model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + sulphatesCent + pHCent + alcoholCent, data=data)
summary(final.model)
```

```{r}
# confidence interval test
confint.default(final.model)
```

We will use this as the final ordianl logistic regression model.

```{r}
# view the final model
kable(tidy(final.model))
```

$log(\frac{P(quality\leq3)}{P(quality>3)})=-6.0844790-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq4)}{P(quality>4)})=-4.1434489-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq5)}{P(quality>5)})=-0.4952119-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq6)}{P(quality>6)})=2.3184245-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq7)}{P(quality>7)})=5.3309018-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$

```{r}
# get CI
ci <- confint(final.model)
```

```{r}
# display odds ratio and CI
exp(cbind(OR = coef(final.model), ci))
```

For one unit increase in volatile.acidity, the odds of the wine falling at or below quality i multiply by a factor of 0.027359812, given that all other variables are held constant.

For one unit increase in citric.acid, the odds of the wine falling at or below quality i multiply by a factor of 0.477221357, given that all other variables are held constant.

For one unit increase in chlorides, the odds of the wine falling at or below quality i multiply by a factor of 0.005199004, given that all other variables are held constant.

For one unit increase in sulphates, the odds of the wine falling at or below quality i multiply by a factor of 15.401872706, given that all other variables are held constant.

For one unit increase in pH, the odds of the wine falling at or below quality i multiply by a factor of 0.228196887, given that all other variables are held constant.

For one unit increase in % alcohol content, the odds of the wine falling at or below quality i multiply by a factor of 2.583998874, given that all other variables are held constant.

### Assumptions
```{r}
# calculate predicted probabilities for each observation
predprobs <- data.frame(predict(final.model,type="probs"))
predprobs %>% filter(row_number() <= 3) # print predictions for first three observations
```

```{r}
# predicted categories
data <- data %>% mutate(pred.comp = predict(final.model,type="class")) # get predicted categories
result <- data %>% group_by(pred.comp,quality_dup) %>% summarise(n=n()) # summarise actual vs. predicted
kable(spread(result,quality_dup,n),format="markdown") # print actual vs. predicted categories
```

### Model Diagnostics
```{r}
data <- data %>% mutate(resid3 = qua3-predprobs[,1],
                        resid4 = qua4-predprobs[,2],
                        resid5 = qua5-predprobs[,3],
                        resid6 = qua6-predprobs[,4],
                        resid7 = qua7-predprobs[,5],
                        resid8 = qua8-predprobs[,6])
data$predor <- predict(final.model, data)
```

```{r}
data %>% count(predor,quality_dup)
confusionMatrix(data$predor,data$quality_dup)
```

```{r}
# binned residual plots of volatile.acidityCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid3,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid4,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid5,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid6,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid7,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid8,xlab="volatile.acidityCent")
```

```{r}
# binned residual plots of citric.acid for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$citric.acid,y=data$resid3,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid4,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid5,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid6,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid7,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid8,xlab="citric.acid")
```

```{r}
# binned residual plots of chloridesCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$chloridesCent,y=data$resid3,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid4,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid5,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid6,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid7,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid8,xlab="chloridesCent")
```

The binned residual plots show a possible chloridesCent outlier.

```{r}
# binned residual plots of sulphatesCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$sulphatesCent,y=data$resid3,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid4,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid5,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid6,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid7,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid8,xlab="sulphatesCent")
```

The binned residual plot shows some potential linear trend and outliers in three categories of quality. We might need to transform sulphatesCent.

```{r}
# binned residual plots of pHCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$pHCent,y=data$resid3,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid4,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid5,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid6,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid7,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid8,xlab="pHCent")
```

There is also some linear trends in the binned residual plots for pHCent. We should also tranform this variable.

```{r}
# binned residual plots of alcohol.Cent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$alcoholCent,y=data$resid3,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid4,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid5,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid6,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid7,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid8,xlab="alcoholCent")
```



```{r}
kable(tidy(vif(final.model)))
```


```{r}
# obtain p-values
tidy(coeftest(final.model))
```

p-values of all the variables are extremely small, so they are all significant predictors of the log-odds of the wine falling in or below quality j. We should also not extrapolate beyond quality j=3,4,...,7.

