---
title: "Project"
author: "Bob Ding, Lynn Fan, Alice Jiang"
date: "11/18/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("cowplot")
library("dplyr")
library("ggplot2")
library("broom")
library("knitr")
library("readr")
library("arm")
library("olsrr")
library("dplyr")
library("car")
library("pROC")
#library("tibble")
```

# Data Exploration - Alice
```{r}
# input data source
data <- read.csv("./redwine_quality.csv")
```

```{r}
# convert quality into factor variable
data <- data %>% mutate(quality = as.factor(quality))
```

```{r}
# overview of the dataset
glimpse(data)
```

```{r}
# scatter plot matrix one
pairs(quality ~ fixed.acidity + volatile.acidity + citric.acid + density + residual.sugar + chlorides, data=data)
``` 

The scatter plot matrix suggests a fairly strong positive linear relationship between fixed.acidity and citric.acid; fixed.acidity and density; and density and residual.sugar. There also appears to be a strong negative linear relationship between volatile.acidity and citric.acid. 

```{r}
# correlation coefficients for potential multicollinearity
cor(data$fixed.acidity, data$citric.acid)
cor(data$fixed.acidity, data$density)
cor(data$volatile.acidity, data$citric.acid)
cor(data$density, data$residual.sugar)
```

The correlation coefficent for fixed.acidity and citric.acid (0.6717034) and the correlation coefficient for fixed.acidity and density (0.6680473) show moderate to strong linear relationship between the variables. To remove potential multicollinearity impact, we **will not include fixed.acidity** in the model.

```{r}
# scatter plot matrix two
pairs(quality ~ free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data=data)
```

The scatter plot shows strong positive linear relationship between free.sulfur.dioxide and total.sulfur.dioxide.

```{r}
# correlation coefficients for potential multicollinearity
cor(data$free.sulfur.dioxide, data$total.sulfur.dioxide)
```

The correlation coefficient of 0.6676665 between free.sulfur.dioxide and total.sulfur.dioxide indicates strong positive linear relationship. We also know from the data description that total.sulfur dioxide includes free.sulfur.dioxide - whose concentration level could impact the smell and taste of the wine, thus impacting quality. So it is sufficient to include only one of these two variables in the model, and we **will include free.sulphur.dioxide**.

```{r}
# distribution of each variable
p1  <- ggplot(data = data, aes(x = quality) ) + geom_histogram(stat="count",fill= "lightblue") 
p3  <- ggplot(data = data, aes(x = volatile.acidity) ) + theme(axis.text=element_text(size=9)) + geom_histogram(fill= "lightblue")
p4  <- ggplot(data = data, aes(x = citric.acid) ) + theme(axis.text=element_text(size=7)) + geom_histogram(fill= "lightblue")
p5  <- ggplot(data = data, aes(x = residual.sugar) ) + geom_histogram(fill= "lightblue")
p6  <- ggplot(data = data, aes(x = chlorides) ) + theme(axis.text=element_text(size=11)) + geom_histogram(fill= "lightblue")
p7  <- ggplot(data = data, aes(x = free.sulfur.dioxide) ) + theme(axis.text=element_text(size=9)) + geom_histogram(fill= "lightblue")
p9  <- ggplot(data = data, aes(x = density) ) + theme(axis.text=element_text(size=7.5)) + geom_histogram(fill= "lightblue")
p10 <- ggplot(data = data, aes(x = pH) ) + geom_histogram(fill= "lightblue")
p11 <- ggplot(data = data, aes(x = sulphates) ) + theme(axis.text=element_text(size=9)) + geom_histogram(fill= "lightblue")
p12 <- ggplot(data = data, aes(x = alcohol) ) + geom_histogram(fill= "lightblue")

plot_grid(p1,p3,p4,p5,p6,p7,p9,p10,p11,p12,ncol = 4,nrow = 3)
```

We can see that free.sulfur.dioxide and alcohol have an obvious rightly skewed distribution. residual.sugar, cholorides, and sulphates also have a slightly rightward skewedness. citric.acid at first appears to have a bimodal distribution, but this is because there are some wines with zero citric acid, so we see a spike at 0 in the histogram. Based on the data definition, we know it is possible for wines to have citric acid of 0. ***** The distribution of citric acid is overall fairly symmetric. ***** All other variables: quality, volatile.acidity, density, and pH are unimodal and fairly symmetric.

******sort quality********

```{r}
# summary statistics of each variable
summary(data)
```

From the previous histogram, we can see that quality has a fairly normal distribution. The summary statistic also shows that quality rating of 5 has the most number of observations (it is also the middle rating on the scale of 0-10). So quality rating of 5 is the most appropriate **reference level** for the model.

```{r}
# makes quality==5 the reference level
data <- data %>% mutate(quality = relevel(quality, ref=5))
```

```{r}
# distribution of citric acid
ggplot(aes(x=citric.acid), data=data) + 
  geom_histogram(fill="steelblue2", color="black") + 
  ggtitle("Distribution of Citric Acid") + 
  theme(plot.title=element_text(color="black", size=14, face="bold.italic", hjust=0.5))
```

```{r}
# add mean-centered numerical variables besides citric.acid
data <- data %>% 
  mutate(volatile.acidityCent = volatile.acidity-mean(volatile.acidity), 
         residual.sugarCent = residual.sugar-mean(residual.sugar), 
         chloridesCent = chlorides-mean(chlorides), 
         free.sulfur.dioxideCent = free.sulfur.dioxide-mean(free.sulfur.dioxide), 
         sulphatesCent = sulphates-mean(sulphates), 
         densityCent = density-mean(density), pHCent = pH-mean(pH),   
         alcoholCent = alcohol-mean(alcohol))
```

# Multiple Regression Analysis
## Lynn 
```{r}
# change quality into a numerical variable
data <- data %>% mutate(quality = as.numeric(quality))
```

```{r}
# primary model
full.model <- lm(quality ~ volatile.acidityCent + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent, data=data)
```

```{r}
# view the primary model
kable(tidy(full.model), format="markdown")
# check model fit
glance(full.model)
```

This is the primary model we constructed based on the exploratory analysis. We can see that volatile.acidityCent, chloridesCent, and pHCent have p-values greater than 0.05. The model's $R^2$ value 0.0559 is noticeably higher than the Adjusted $R^2$ of 0.0506, penalizing for unnecessary variables in the model. 

We will then use model selection to build a better model, first including only the main effects.

## Model Selections
```{r}
# backward selection process
backward <- ols_step_backward_aic(full.model, details=TRUE)
```

```{r}
# view stepwise AIC backward elimination
plot(backward)
```

The backward selection removes pHCent and volatile.acidityCent from the model.

```{r}
# forward selection process
forward <- ols_step_forward_aic(full.model, details=TRUE)
```

```{r}
# view stepwise AIC forward selection
plot(forward)
```

We can see the foward selection process includes the same explanatory variables as the model from the backward selection - both models do not include volatile.acidityCent and pHCent from the full.model.

```{r}
# stepwise selection process
stepwise <- ols_step_both_aic(full.model, details=TRUE)
```

```{r}
# view stepwise AIC two-direction selection
plot(stepwise)
``` 

The plot for stepwise selection looks the same as the plot for forward selection. Therefore, for all three model selections, we obtain the same result of including all variables from the full.model, except for volatile.acidityCent and pHCent.

```{r}
# new model from model selection
selected <- lm(quality ~ residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent, data=data)
```

```{r}
# view coefficients of the new model
kable(tidy(selected), format="markdown")
```

We can see that chloridesCent has a p-value much greater than 0.05, so we consider chloridesCent to be a statistically insignificant predictor and remove it from the model. 

```{r}
# model after removing chloridesCent
model <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent, data=data)
```

```{r}
# view the model
kable(tidy(model), format="markdown")
```

We can see that the p-values of most of the explanatory variables are smaller than the threshold 0.05, which means there is sufficient evidence that these varaibles are significant predictors of wine quality. There is one exception, which is densityCent, whose p-value of 0.06 is slightly greater than 0.05. We remember that the p-value of densityCent is the selected model that included chloridesCent was 0.45 (already very close to 0.05). We believe 0.06 p-value is a tolerable level and will still keep densityCent in the model, since it helps with fitting the model.

```{r}
# multicollinearity check
kable(tidy(vif(model)))
```

A quick examination of the model using the VIF() function shows no noticeable detection of multicollinearity effects - all vif values are quite small, ranging around 1 to 2.

We will then examine possible interaction effects between the selected variables.

## Nested F Tests on Interactions
```{r}
# interaction effects with residual.sugarCent
model_rs <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + residual.sugarCent*citric.acid + residual.sugarCent*free.sulfur.dioxideCent + residual.sugarCent*sulphatesCent + residual.sugarCent*densityCent + residual.sugarCent*alcoholCent, data=data)
```

```{r}
# Nested F Test for interaction with residual.sugarCent
ano1 <- anova(model_rs, model)
kable(tidy(ano1))
```

p-value is much higher than 0.05, so there is no significant interactions with residual.sugarCent that impact wine quality.

```{r}
# Nested F Test for interaction effect between alcohol and density
model_da <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + alcoholCent*densityCent, data=data)
ano2 <- anova(model_da, model)
kable(tidy(ano2))
```

p-value is much larger than 0.05, so there is no significant interaction effect between alcoholCent an densityCent. 

```{r}
# Nested F Test for all other two-way interactions
model1 <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + citric.acid*densityCent + citric.acid*alcoholCent + free.sulfur.dioxideCent*sulphatesCent + free.sulfur.dioxideCent*densityCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*densityCent + sulphatesCent*alcoholCent, data=data)
ano3 <- anova(model1, model)
kable(tidy(ano3))
```

The p-value is extremely small, so there is at least one significant interaction effects. We will then investigate which specific interactions are significant using backward model selection.

```{r}
# backward selection
backward_new <- ols_step_backward_aic(model1, details=TRUE)
```

```{r}
# view the stepwise AIC backward elimination
plot(backward_new)
```

Backward model selection removes 5 interaction effects and sulphatesCent. However, since the model still includes some interaction effects with sulphatesCent, we will keep this main effect in the model.

```{r}
# new model after backward selection
model2 <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*alcoholCent, data=data)
kable(tidy(model2), format="markdown")
```

We can see that p-value of all the coefficients are less than the threshold of 0.05, except for sulphatesCent. Therefore, these main and interaction effects are significant predictors of wine quality (besides sulphatesCent). So we will make model2 our final model.

```{r}
# make model2 the final multiple regression model
final.model <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*alcoholCent, data=data)
```

Our final multiple reression model:  
$quality = 4.2245 - 0.0512*residual.sugarCent - 0.6498*citric.acid + 0.0204*free.sulfur.dioxideCent - 0.1836*sulphatesCent + 57.4648*densityCent - 0.1355*alcoholCent - 0.0438*citric.acid*free.sulfur.dioxideCent - 1.9995*citric.acid*sulphatesCent + 0.0103*free.sulfur.dioxideCent*alcoholCent - 1.3477*sulphatesCent*alcoholCent$  

## Assumptions

### Residual Plots
```{r}
data <- data %>% mutate(predicted = predict.lm(final.model))
data <- data %>% mutate(resid=resid(final.model))
ggplot(data, aes(predicted, resid)) + geom_point() + 
  geom_hline(yintercept=0, color="red") + labs(x="Predicted Values", y="Residuals", 
  title="Residuals vs. Fitted Values") + theme(plot.title=element_text(hjust=0.5, 
  size=10))
```

### Normaltiy
```{r}
# distribution of residuals
ggplot(data, aes(resid)) + geom_histogram(color="black", 
  fill="steelblue2") + labs(x="residuals", y="count", title="Histogram of Residuals") + 
  theme(plot.title=element_text(hjust=0.5, size=10))
# QQ Plot
qqnorm(bikeshare$resid, main="Normal QQ Plot of Residuals")
qqline(bikeshare$resid, col="red")
```

The histogram is unimodal and fairly symmetric. In the QQ plot, we can see most of the points lie right on top of the red diagonal line, with some deviation on the sides. Overall, the Normaltiy Assumption is well satisfied.

### Influence Points
```{r}
data <- data %>% mutate(leverage = hatvalues(final.model), cooks = 
  cooks.distance(final.model), stand.resid = rstandard(final.model), obs.num =  
  row_number())
```

```{r}
# calculate leverage threshold
t <- 2*10/1599
```

```{r}
# leverage
ggplot(data=data, aes(x=obs.num, y=leverage)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept=t, color="red") +
  labs(x="Observation Number", y="Leverage", title="Leverage")
```

We can see that there is one point with a significantly high leverage around 0.4, comparing with other observations. This could be an outlier and might be an influence point.

```{r}
data %>% filter(leverage > 0.3) %>% 
  select(citric.acid, residual.sugarCent, free.sulfur.dioxideCent, sulphatesCent, densityCent, alcoholCent)
```

```{r}
# Cook’s Distance
ggplot(data=data, aes(x=obs.num, y=cooks)) + 
  geom_point() + 
  geom_hline(yintercept=1, color="red")+
  labs(x="Observation Number", y="Cook's Distance", title="Cook's Distance")
```

The Cook’s Distance for all observations are far below the threshold of 1.

```{r}
# standardized residuals
p1 <- ggplot(data=data, aes(x=predicted, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p2 <- ggplot(data=data, aes(x=residual.sugarCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p3 <- ggplot(data=data, aes(x=citric.acid, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p4 <- ggplot(data=data, aes(x=free.sulfur.dioxideCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p5 <- ggplot(data=data, aes(x=sulphatesCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p6 <- ggplot(data=data, aes(x=densityCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p7 <- ggplot(data=data, aes(x=alcoholCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")

plot_grid(p1, p2, p3, p4, p5, p6, p7, ncol=2, nrow=4)
```

The standardized residuals show some points with magnitude greater than 2, but overall, in combination with our observation from Cook’s Distance and just one data point with high leverage away from other points, we can conclude there isn't any obvious influential points in this model.

<!--citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*alcoholCent-->

### Multicollinearity
```{r}
kable(tidy(vif(final.model)))
```

## Aditional Work
```{r}
model_fsd <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + free.sulfur.dioxideCent*sulphatesCent + free.sulfur.dioxideCent*densityCent + free.sulfur.dioxideCent*alcoholCent, data=data)
ano3 <- anova(model_fsd, model)
kable(tidy(ano3))
```

```{r}
model_s <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + sulphates*densityCent + sulphatesCent*alcoholCent, data=data)
ano4 <- anova(model_s, model)
kable(tidy(ano4))
```

# check using stepwise
```{r}
# stepwise selection
stepwise_new <- ols_step_both_aic(model1, details=TRUE)
# view the stepwise AIC backward elimination
plot(stepwise_new)
```

# Bob Logistic Regression

```{r}
rawdata <- read.csv("./redwine_quality.csv")
data$quality <- rawdata$quality
data$quality[data$quality < 6] <- 0
data$quality[data$quality >= 6] <- 1

# Model Reduced
modellg1.null <- glm(quality ~ 1, data=data , family=binomial) 

# Model Full
modellg1.full <- glm(quality ~ volatile.acidityCent  + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent,data=data, family = binomial)
```

```{r}
# forward Selection
step(modellg1.null,scope=list(upper=modellg1.full),direction="forward")
```

```{r}
modellg1.selected <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + pHCent + densityCent,data=data, family = binomial)
summary(modellg1.selected)
kable(tidy(modellg1.selected))
```

```{r}
modellg1.final <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid,data=data, family = binomial)
summary(modellg1.final)
kable(tidy(modellg1.final))
```

```{r}
kable(tidy(vif(modellg1.final)))
```

```{r}
data$pred <- predict.glm(modellg1.final)
data$res <- residuals.glm(modellg1.final)
```

```{r,fig.height=4,fig.width=4,echo=F}
binnedplot(x=data$pred,y=data$res,xlab="Predict")
binnedplot(x=data$alcoholCent,y=data$res,xlab="alcoholCent")
binnedplot(x=data$volatile.acidityCent,y=data$res,xlab="volatile.acidityCent")
binnedplot(x=data$sulphatesCent,y=data$res,xlab="sulphatesCent")
binnedplot(x=data$chloridesCent,y=data$res,xlab="chloridesCent")
binnedplot(x=data$free.sulfur.dioxideCent,y=data$res,xlab="free.sulfur.dioxideCent")
binnedplot(x=data$citric.acid,y=data$res,xlab="citric.acid")
```

sulfates cent need transformation
volatile.acidityCent also needs tranformation
alcoholCent also needs transformation

```{r}
avPlots(modellg1.final)
cutoff <- cooks.distance(modellg1.final)
plot(modellg1.final, which=4, cook.levels=cutoff)
influencePlot(modellg1.final,	 main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )

data <- data %>%
  mutate(leverage = hatvalues(modellg1.final), 
         cooks = cooks.distance(modellg1.final),
         stand.resid = rstandard(modellg1.final), 
         obs.num = row_number())
p3 <- ggplot(data=data, aes(x=obs.num,y=leverage)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept=0.1,color="red")+
  labs(x="Observation Number",y="Leverage",title="Leverage")

p4 <- ggplot(data=data, aes(x=obs.num,y=cooks)) + 
  geom_point() + 
  geom_hline(yintercept=1,color="red")+
  labs(x="Observation Number",y="Cook's Distance",title="Cook's Distance")

plot_grid(p3,p4,ncol = 2)
```

no significant outlier or influential plot.

## Lynn-Interactions on Logistic Regression

```{r}
# interaction effects with change in deviance test
modellg1.final <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid,data=data, family = binomial)
modellg2 <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + alcoholCent*volatile.acidityCent, data=data, family = binomial)
modellg3 <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + alcoholCent*sulphatesCent, data=data, family = binomial)
modellg4 <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + volatile.acidityCent*sulphatesCent, data=data, family = binomial)
tidy(anova(modellg1.final,modellg2,test="Chisq"))
tidy(anova(modellg1.final,modellg3,test="Chisq"))
tidy(anova(modellg1.final,modellg4,test="Chisq"))
```
Based on the three test, we discover significant interaction between alcoholCent and sulphatesCent, and we are including it in our model.


```{r}
#final logistic regression model
final.model.lg <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + alcoholCent*sulphatesCent, data=data, family = binomial)
summary(final.model.lg)
```

Our final logistic regression model:  
$quality = 0.539269 + 0.996689*alcoholCent - 3.519343*volatile.acidityCent + 3.367093*sulphatesCent - 2.259180*chloridesCent - 0.014257*free.sulfur.dioxideCent - 0.976473*citric.acid + 1.549278*alcoholCent*sulphatesCent $ 

## Ordinal
