---
title: "Project"
author: "Bob Ding, Lynn Fan, Alice Jiang"
date: "11/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("cowplot")
library("dplyr")
library("ggplot2")
library("broom")
library("knitr")
library("readr")
library("arm")
library("olsrr")
library("dplyr")
library("car")
library("pROC")
library("tibble")
library("nnet")
library("tidyr")
library("MASS")
library("caret")
library("AER")
```

# Data Exploration
```{r}
# input data source
data <- read.csv("./redwine_quality.csv")
```

```{r}
# overview of the dataset
glimpse(data)
```

```{r}
# convert quality into factor variable
data <- data %>% mutate(quality = as.factor(quality))
```

```{r}
# duplicate quality for ordinal regresison
data <- data %>% mutate(quality_dup = as.factor(quality))
```

<!--
```{r}
# overview of the dataset
kable(tidy(summary(data)))
```
-->

```{r}
# scatter plot matrix one
pairs(quality ~ fixed.acidity + volatile.acidity + citric.acid + density + residual.sugar + chlorides, data=data)
``` 

The scatter plot matrix suggests a fairly strong positive linear relationship between fixed.acidity and citric.acid; fixed.acidity and density; and density and residual.sugar. There also appears to be a strong negative linear relationship between volatile.acidity and citric.acid. 

```{r}
# correlation coefficients for potential multicollinearity
cor(data$fixed.acidity, data$citric.acid)
cor(data$fixed.acidity, data$density)
cor(data$volatile.acidity, data$citric.acid)
cor(data$density, data$residual.sugar)
```

The correlation coefficent for fixed.acidity and citric.acid (0.6717034) and the correlation coefficient for fixed.acidity and density (0.6680473) show moderate to strong linear relationship between the variables. To remove potential multicollinearity impact, we **will not include fixed.acidity** in the model.

```{r}
# scatter plot matrix two
pairs(quality ~ free.sulfur.dioxide + total.sulfur.dioxide + pH + sulphates + alcohol, data=data)
```

The scatter plot shows strong positive linear relationship between free.sulfur.dioxide and total.sulfur.dioxide.

```{r}
# correlation coefficients for potential multicollinearity
cor(data$free.sulfur.dioxide, data$total.sulfur.dioxide)
```

The correlation coefficient of 0.6676665 between free.sulfur.dioxide and total.sulfur.dioxide indicates strong positive linear relationship. We also know from the data description that total.sulfur dioxide includes free.sulfur.dioxide - whose concentration level could impact the smell and taste of the wine, thus impacting quality. So it is sufficient to include only one of these two variables in the model, and we **will include free.sulphur.dioxide**.

```{r}
# distribution of each variable
p1  <- ggplot(data = data, aes(x = quality) ) + geom_histogram(stat="count",fill= "lightblue") 
p2  <- ggplot(data = data, aes(x = volatile.acidity) ) + theme(axis.text=element_text(size=9)) + geom_histogram(fill= "lightblue")
p3  <- ggplot(data = data, aes(x = citric.acid) ) + theme(axis.text=element_text(size=7)) + geom_histogram(fill= "lightblue")
p4  <- ggplot(data = data, aes(x = residual.sugar) ) + geom_histogram(fill= "lightblue")
p5  <- ggplot(data = data, aes(x = chlorides) ) + theme(axis.text=element_text(size=11)) + geom_histogram(fill= "lightblue")
p6  <- ggplot(data = data, aes(x = free.sulfur.dioxide) ) + theme(axis.text=element_text(size=9)) + geom_histogram(fill= "lightblue")
p7  <- ggplot(data = data, aes(x = density) ) + theme(axis.text=element_text(size=7.5)) + geom_histogram(fill= "lightblue")
p8 <- ggplot(data = data, aes(x = pH) ) + geom_histogram(fill= "lightblue")
p9 <- ggplot(data = data, aes(x = sulphates) ) + theme(axis.text=element_text(size=9)) + geom_histogram(fill= "lightblue")
p10 <- ggplot(data = data, aes(x = alcohol) ) + geom_histogram(fill= "lightblue")

plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,ncol=4,nrow=3)
```

We can see that free.sulfur.dioxide and alcohol have an obvious rightly skewed distribution. residual.sugar, cholorides, and sulphates also have a slightly rightward skewedness. citric.acid appears to have a bimodal distribution; however, it could be due to the fact that some wines are allowed to have zero citric acid, so we see a spike at 0 in the histogram. All other variables: quality, volatile.acidity, density, and pH are unimodal and fairly symmetric. <!--We will consider transformation on free.sulfur.dioxide and alcohol. -->

```{r}
# summary statistics of each variable
summary(data)
```

From the previous histogram, we can see that quality has a fairly normal distribution. The summary statistic also shows that quality rating of 5 has the most number of observations (it is also the middle rating on the scale of 0-10). So quality rating of 5 is the most appropriate **reference level** for the model.

```{r}
# makes quality==5 the reference level
data <- data %>% mutate(quality = relevel(quality, ref=5))
```

<!--
```{r}
# distribution of citric acid
ggplot(aes(x=citric.acid), data=data) + 
  geom_histogram(fill="steelblue2", color="black") + 
  ggtitle("Distribution of Citric Acid") + 
  theme(plot.title=element_text(color="black", size=14, face="bold.italic", hjust=0.5))
```
-->

```{r}
# add mean-centered numerical variables besides citric.acid
data <- data %>% 
  mutate(volatile.acidityCent = volatile.acidity-mean(volatile.acidity), 
         residual.sugarCent = residual.sugar-mean(residual.sugar), 
         chloridesCent = chlorides-mean(chlorides), 
         free.sulfur.dioxideCent = free.sulfur.dioxide-mean(free.sulfur.dioxide), 
         sulphatesCent = sulphates-mean(sulphates), 
         densityCent = density-mean(density), pHCent = pH-mean(pH),   
         alcoholCent = alcohol-mean(alcohol))
```

# Multiple Regression Analysis
```{r}
# change quality into a numerical variable
data <- data %>% mutate(quality = as.numeric(quality))
```

```{r}
# primary model
full.model <- lm(quality ~ volatile.acidityCent + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent, data=data)
```

```{r}
# view the primary model
kable(tidy(full.model), format="markdown")
# check model fit
glance(full.model)
```

This is the primary model we constructed based on the exploratory analysis. We can see that volatile.acidityCent, chloridesCent, and pHCent have p-values greater than 0.05. The model's $R^2$ value 0.05593565 is higher than the Adjusted $R^2$ of 0.05058853, penalizing for unnecessary variables in the model. 

We will then use model selection to build a better model, first including only the main effects.

## Model Selections
```{r}
# backward selection process
backward <- ols_step_backward_aic(full.model, details=TRUE)
```

```{r}
# view stepwise AIC backward elimination
plot(backward)
```

The backward selection removes pHCent and volatile.acidityCent from the model.

```{r}
# forward selection process
forward <- ols_step_forward_aic(full.model, details=TRUE)
```

```{r}
# view stepwise AIC forward selection
plot(forward)
```

We can see the foward selection process includes the same explanatory variables as the model from the backward selection - both models do not include volatile.acidityCent and pHCent from the full.model.

```{r}
# stepwise selection process
stepwise <- ols_step_both_aic(full.model, details=TRUE)
```

```{r}
# view stepwise AIC two-direction selection
plot(stepwise)
``` 

The plot for stepwise selection looks the same as the plot for forward selection. Therefore, for all three model selections, we obtain the same result of including all variables from the full.model, except for volatile.acidityCent and pHCent.

```{r}
# new model from model selection
selected <- lm(quality ~ residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent, data=data)
```

```{r}
# view coefficients of the new model
kable(tidy(selected), format="markdown")
```

We can see that chloridesCent has a p-value of 0.0911197, which is much greater than 0.05. So we consider chloridesCent to be a statistically insignificant predictor and remove it from the model. 

```{r}
# model after removing chloridesCent
model <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent, data=data)
```

```{r}
# view the model
kable(tidy(model), format="markdown")
```

We can see that the p-values of most of the explanatory variables are smaller than the threshold 0.05, which means there is sufficient evidence that these varaibles are significant predictors of wine quality. There is one exception, which is densityCent, whose p-value of 0.0614575 is slightly greater than 0.05. We remember that the p-value of densityCent is the selected model that included chloridesCent was 0.0450413 (already very close to 0.05). We believe 0.06 p-value is a tolerable level and will still keep densityCent in the model, since it helps with fitting the model.

```{r}
# multicollinearity check
kable(tidy(vif(model)))
```

A quick examination of the model using the VIF() function shows no noticeable detection of multicollinearity effects - all vif values are quite small, ranging around 1 to 2.

We will then examine possible interaction effects between the selected variables.

## Nested F Tests on Interactions
```{r}
# interaction effects with residual.sugarCent
model_rs <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + residual.sugarCent*citric.acid + residual.sugarCent*free.sulfur.dioxideCent + residual.sugarCent*sulphatesCent + residual.sugarCent*densityCent + residual.sugarCent*alcoholCent, data=data)
```

```{r}
# Nested F Test for interaction with residual.sugarCent
ano1 <- anova(model_rs, model)
kable(tidy(ano1))
```

p-value is much higher than 0.05, so there is no significant interactions with residual.sugarCent that impact wine quality.

```{r}
# Nested F Test for interaction effect between alcohol and density
model_da <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + alcoholCent*densityCent, data=data)
ano2 <- anova(model_da, model)
kable(tidy(ano2))
```

p-value is much larger than 0.05, so there is no significant interaction effect between alcoholCent an densityCent. 

```{r}
# Nested F Test for all other two-way interactions
model1 <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + citric.acid*densityCent + citric.acid*alcoholCent + free.sulfur.dioxideCent*sulphatesCent + free.sulfur.dioxideCent*densityCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*densityCent + sulphatesCent*alcoholCent, data=data)
ano3 <- anova(model1, model)
kable(tidy(ano3))
```

The p-value is extremely small, so there is at least one significant interaction effects. We will then investigate which specific interactions are significant using backward model selection.

```{r}
# backward selection
backward_new <- ols_step_backward_aic(model1, details=TRUE)
```

```{r}
# view the stepwise AIC backward elimination
plot(backward_new)
```

Backward model selection removes 5 interaction effects and sulphatesCent. However, since the model still includes some interaction effects with sulphatesCent, we will keep this main effect in the model.

```{r}
# new model after backward selection
model2 <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*alcoholCent, data=data)
kable(tidy(model2), format="markdown")
```

We can see that p-value of all the coefficients are less than the threshold of 0.05, except for sulphatesCent. Therefore, these main and interaction effects are significant predictors of wine quality (besides sulphatesCent). So we will make model2 our final model.

```{r}
# make model2 the final multiple regression model
final.model <- lm(quality ~ residual.sugarCent + citric.acid + free.sulfur.dioxideCent + sulphatesCent + densityCent + alcoholCent + citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*alcoholCent, data=data)
```

Our final multiple reression model:  
$quality = 4.2245 - 0.0512*residual.sugarCent - 0.6498*citric.acid + 0.0204*free.sulfur.dioxideCent - 0.1836*sulphatesCent + 57.4648*densityCent - 0.1355*alcoholCent - 0.0438*citric.acid*free.sulfur.dioxideCent - 1.9995*citric.acid*sulphatesCent + 0.0103*free.sulfur.dioxideCent*alcoholCent - 1.3477*sulphatesCent*alcoholCent$  

## Assumptions

### Residual Plots
```{r}
data <- data %>% mutate(predicted = predict.lm(final.model))
data <- data %>% mutate(resid=resid(final.model))
ggplot(data, aes(predicted, resid)) + geom_point() + 
  geom_hline(yintercept=0, color="red") + labs(x="Predicted Values", y="Residuals", 
  title="Residuals vs. Fitted Values") + theme(plot.title=element_text(hjust=0.5, 
  size=10))
```

This residual plot is a result of having categorical response variable.

### Normaltiy
```{r}
# distribution of residuals
ggplot(data, aes(resid)) + geom_histogram(color="black", 
  fill="steelblue2") + labs(x="residuals", y="count", title="Histogram of Residuals") + 
  theme(plot.title=element_text(hjust=0.5, size=10))
# QQ Plot
qqnorm(data$resid, main="Normal QQ Plot of Residuals")
qqline(data$resid, col="red")
```
 
The distribution of residuals appear to be bimodal. The QQ plot also suggests the same conclusion, since we can see a very prominent deviation from the diagonal normal line on the left side. Overall, the Normaltiy Assumption seems to be violated. This could be due to the fact that we are dealing with a categorical response varaible quality.

### Influence Points
```{r}
data <- data %>% mutate(leverage = hatvalues(final.model), cooks = 
  cooks.distance(final.model), stand.resid = rstandard(final.model), obs.num =  
  row_number())
```

```{r}
# calculate leverage threshold
t <- 2*10/1599
```

```{r}
# leverage
ggplot(data=data, aes(x=obs.num, y=leverage)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept=t, color="red") +
  labs(x="Observation Number", y="Leverage", title="Leverage")
```

We can see that there is one point with a significantly high leverage around 0.4, comparing with other observations. This could be an outlier and might be an influence point.

```{r}
#data %>% filter(leverage > 0.3) %>% 
  #select(citric.acid, residual.sugarCent, free.sulfur.dioxideCent, sulphatesCent, densityCent, alcoholCent)
```

```{r}
# Cook’s Distance
ggplot(data=data, aes(x=obs.num, y=cooks)) + 
  geom_point() + 
  geom_hline(yintercept=1, color="red")+
  labs(x="Observation Number", y="Cook's Distance", title="Cook's Distance")
```

The Cook’s Distance for all observations are far below the threshold of 1.

```{r}
# standardized residuals
p1 <- ggplot(data=data, aes(x=predicted, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p2 <- ggplot(data=data, aes(x=residual.sugarCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p3 <- ggplot(data=data, aes(x=citric.acid, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p4 <- ggplot(data=data, aes(x=free.sulfur.dioxideCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p5 <- ggplot(data=data, aes(x=sulphatesCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p6 <- ggplot(data=data, aes(x=densityCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")
p7 <- ggplot(data=data, aes(x=alcoholCent, y=stand.resid)) + geom_point()+  geom_hline(yintercept=0, color="red")

plot_grid(p1, p2, p3, p4, p5, p6, p7, ncol=2, nrow=4)
```

The standardized residuals show some points with magnitude greater than 2, but overall, in combination with our observation from Cook’s Distance and just one data point with high leverage away from other points, we can conclude there isn't any obvious influential points in this model.

<!--citric.acid*free.sulfur.dioxideCent + citric.acid*sulphatesCent + free.sulfur.dioxideCent*alcoholCent + sulphatesCent*alcoholCent-->

### Multicollinearity
```{r}
kable(tidy(vif(final.model)))
```

Overall, the final model does not have any major multicollinearity concerns - all VIF values are fairly low and below the threshold of 10. However, multiple regression model does not seem to be the best fitting for our data set. Given we have a categorical response variable, we will try different types of logistic regresison models.

# Logistic Regression

```{r}
# transform quality into a binary variable
rawdata <- read.csv("./redwine_quality.csv")
data$quality <- rawdata$quality
data$quality[data$quality < 6] <- 0    # any quality less than 6 is 0 (not good)
data$quality[data$quality >= 6] <- 1   # quality equal to or greater than 6 so 6-10 is 1 (good)
```

```{r}
# Model Reduced
modellg1.null <- glm(quality ~ 1, data=data , family=binomial) 

# Model Full
modellg1.full <- glm(quality ~ volatile.acidityCent + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent, data=data, family = binomial)
```

```{r}
# forward Selection
step(modellg1.null,scope=list(upper=modellg1.full),direction="forward")
```

```{r}
# model based on forward selection
modellg1.selected <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + pHCent + densityCent,data=data, family = binomial)
summary(modellg1.selected)
```

```{r}
# view coefficients of the selected model
kable(tidy(modellg1.selected))
```

We can see that the p-value of pHCent 0.0507387 is slightly higher than the 0.05 threshold. densityCent has a p-value 0.0947266 greater than 0.05. We will build a new model without these two variables and analyze the result.

```{r}
# model after removing pHCent and densityCent
modellg1.final <- glm(quality ~ alcoholCent + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid, data=data, family = binomial)
summary(modellg1.final)
```

```{r}
# view coefficients of the new model
kable(tidy(modellg1.final))
```

All the variables have p-values much smaller than 0.05, so there is significant evidence that they are important predictors of the log-odds (and therefore odds) of wine quality (good vs. not good). Based on the model, it appears that the percent alcohol content of the wine is the strongest predictor of wine quality. alcoholCent has the largest test statistic magnitude of 13.402476. The positive test statistic value also shows that as the % alcohol content of the wine increases, the logs-odds of good vs. not good wine quality increases. The amount of acetic acid in wine (volatile.acidityCent) and the amount of wine additive (sulphatesCent) are also strong predictors.

## Assumptions
```{r}
# calculate predicted probabilities
data$pred <- predict.glm(modellg1.final, type="response")
# calculate residuals of predicted probabilities
data$res <- residuals.glm(modellg1.final, type="response")
```

```{r,fig.height=4,fig.width=4,echo=F}
# binned residual plots
binnedplot(x=data$pred,y=data$res,xlab="Predict")
binnedplot(x=data$alcoholCent,y=data$res,xlab="alcoholCent")
binnedplot(x=data$volatile.acidityCent,y=data$res,xlab="volatile.acidityCent")
binnedplot(x=data$sulphatesCent,y=data$res,xlab="sulphatesCent")
binnedplot(x=data$chloridesCent,y=data$res,xlab="chloridesCent")
binnedplot(x=data$free.sulfur.dioxideCent,y=data$res,xlab="free.sulfur.dioxideCent")
binnedplot(x=data$citric.acid,y=data$res,xlab="citric.acid")
```

Most of the binned residual plots show random pattern and raises no major concerns of violations. The binned residual plot of alcoholCent and free.sulfur.dioxideCent seems to show an outlier on the top left, suggesting high average residuals for that bin. There is also one outlier on the binned residual plot of chloridesCent on the far right. 

The binned residual plot also reveals that sulphatesCent might need a transformation, since there appears to be some linear trend.

```{r}
# use AV Plot to identify prblematic observations
avPlots(modellg1.final)
```

We can see observations 653 and 93 are consistently identified in av plots. We might consider removing these points if they are influential.

```{r}
# influence points
data <- data %>%
  mutate(leverage = hatvalues(modellg1.final), 
         cooks = cooks.distance(modellg1.final),
         stand.resid = rstandard(modellg1.final), 
         obs.num = row_number())
```

```{r}
# Cook's Distance
cutoff <- cooks.distance(modellg1.final)
plot(modellg1.final, which=4, cook.levels=cutoff)
```

We can see that observation 93 has the greatest Cook's Distance. However, its Cook's Distance is still below 0.05 and is much smaller than the threshold of 1.

```{r}
# influence plot
influencePlot(modellg1.final,	 main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )
```

We can see that some observations have high studentized residual value with magnitude greater than 2, but they have very low hat values and acceptable circle size. Observations 259 and 152 have very high leverage (hat values) but are within reasonable studentized residual value and circle size. Overall, there is no outstanding influence point in the data set, based on considerations of leverage and cook's distance.

```{r}
# leverage threshold
threshold <- 0.05
```

Given we have 1599 observations and the above influence plot, we believe 0.05 is a reasonable threshold for leverage.

```{r}
# leverage and cook's distance plots
p1 <- ggplot(data=data, aes(x=obs.num,y=leverage)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept=threshold,color="red")+
  labs(x="Observation Number",y="Leverage",title="Leverage")

p2 <- ggplot(data=data, aes(x=obs.num,y=cooks)) + 
  geom_point() + 
  geom_hline(yintercept=1,color="red")+
  labs(x="Observation Number",y="Cook's Distance",title="Cook's Distance")

plot_grid(p1,p2,ncol = 2)
```

Overall, there is no significant influence point.

```{r}
# multicollinearity
kable(tidy(vif(modellg1.final)))
```

VIF values are very small, so there is no major concerns of multicollinearity.

```{r}
# ROC Curve
ROC.quality <- roc(data$quality, data$pred, plot=T)
```

```{r}
# area under the curve
ROC.quality$auc
```

From the ROC curve and AUC calculation, we can see the curve is fairly close to the top left corner (area under the curve is close to 1). This shows that the logistic model is able to distinguish between good and not good quality, so this is a pretty good model.

## Interaction Effects in Logistic Regression
```{r}
# interaction effect between alcoholCent and volatile.acidityCent using change in deviance test
modellg1.final <- glm(quality ~ alcoholCent + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid,data=data, family = binomial) # the final logistic regression model
modellg1 <- glm(quality ~ alcoholCent + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + alcoholCent*volatile.acidityCent, data=data, family = binomial)
tidy(anova(modellg1.final, modellg1, test="Chisq"))
```

p-value is greater than 0.05, so this interaction effect is not significant.

```{r}
# interaction effect between alcoholCent and volatile.acidityCent
modellg2 <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + volatile.acidityCent*sulphatesCent, data=data, family = binomial)
tidy(anova(modellg1.final, modellg2, test="Chisq"))
```

This interaction effect is also not significant.

```{r}
# interaction effect between alcoholCent and sulphates
modellg3 <- glm(quality ~ alcoholCent + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + alcoholCent*sulphatesCent, data=data, family = binomial)
tidy(anova(modellg1.final, modellg3, test="Chisq"))
```

p-value is smaller than 0.05, so the interaction effect between alcoholCent and sulphatesCent is significant, and we will include it in our model.

***************
```{r}
# final logistic regression model with interaction effects
final.model.lg <- glm(quality ~ alcoholCent  + volatile.acidityCent + sulphatesCent + chloridesCent + free.sulfur.dioxideCent + citric.acid + alcoholCent*sulphatesCent, data=data, family = binomial)
summary(final.model.lg)
```

```{r}
# view the final model
kable(tidy(final.model.lg))
```
******************p-value of chloridesCent > 0.05...

Our final logistic regression model:  
$\frac{\pi_{good}}{\pi_{bad}} = \exp(0.539269 + 0.996689*alcoholCent - 3.519343*volatile.acidityCent + 3.367093*sulphatesCent - 2.259180*chloridesCent - 0.014257*free.sulfur.dioxideCent - 0.976473*citric.acid + 1.549278*alcoholCent*sulphatesCent)$ 

# Ordinal Logistic Regression
Because the response variable, quality, is ordinal and has a natural ordering for categories 0,1,...,10, we believe ordinal logistic regression might be a good model.

```{r}
# create separate indicator variables for each level of quality to use in binned plots
data <- data %>% mutate(qua0 = if_else(quality_dup=="0",1,0), 
                        qua1 = if_else(quality_dup=="1",1,0),
                        qua2 = if_else(quality_dup=="2",1,0),
                        qua3 = if_else(quality_dup=="3",1,0),
                        qua4 = if_else(quality_dup=="4",1,0),
                        qua5 = if_else(quality_dup=="5",1,0),
                        qua6 = if_else(quality_dup=="6",1,0),
                        qua7 = if_else(quality_dup=="7",1,0),
                        qua8 = if_else(quality_dup=="8",1,0),                            
                        qua9 = if_else(quality_dup=="9",1,0),
                        qua10 = if_else(quality_dup=="10",1,0))
                          
``` 

```{r}
# full model
full.model <- polr(quality_dup ~ volatile.acidityCent + residual.sugarCent + citric.acid + chloridesCent + free.sulfur.dioxideCent + sulphatesCent + densityCent + pHCent + alcoholCent, data=data)
summary(full.model)
```

```{r}
# view the model
kable(tidy(full.model),format="markdown",digits=3)
``` 

We see there are 5 possible intercepts/models. We also observe the magnitude of the test statistic for densityCent is significantly higher than the test statistic magnitude of other variables.

```{r}
# parameter estimates with confidence level
confint.default(full.model)
``` 

If the 95% CI does not cross 0, the parameter estimate is statistically significant. The CI for volatile.acidityCent, citric.acid, chloridesCent, sulphatesCent, densityCent, pHCent and alcoholCent do not include 0, so we take them as significant and remove residual.sugarCent and free.sulfur.dioxideCent from the model.

```{r}
# new model
model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + sulphatesCent + pHCent + densityCent + alcoholCent, data=data)
summary(model)
```

```{r}
# confidence interval test of the new model
confint.default(model)
```

densityCent contains zero in its 95% CI.

```{r}
# new model without densityCent
final.model <- polr(quality_dup ~ volatile.acidityCent + citric.acid + chloridesCent + sulphatesCent + pHCent + alcoholCent, data=data)
summary(final.model)
```

```{r}
# confidence interval test
confint.default(final.model)
```

We will use this as the final ordianl logistic regression model.

```{r}
# view the final model
kable(tidy(final.model))
```

$log(\frac{P(quality\leq3)}{P(quality>3)})=-6.0844790-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq4)}{P(quality>4)})=-4.1434489-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq5)}{P(quality>5)})=-0.4952119-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq6)}{P(quality>6)})=2.3184245-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$
$log(\frac{P(quality\leq7)}{P(quality>7)})=5.3309018-[-3.5986801*volatile.acidityCent - 0.7397748*citric.acid - 5.2592883*chloridesCent + 2.7344891*sulphatesCent	- 1.4775465*pHCent + 0.9493381*alcoholCent]$

```{r}
# get CI
ci <- confint(final.model)
```

```{r}
# display odds ratio and CI
exp(cbind(OR = coef(final.model), ci))
```

For one unit increase in volatile.acidity, the odds of the wine falling at or below quality i multiply by a factor of 0.027359812, given that all other variables are held constant.

For one unit increase in citric.acid, the odds of the wine falling at or below quality i multiply by a factor of 0.477221357, given that all other variables are held constant.

For one unit increase in chlorides, the odds of the wine falling at or below quality i multiply by a factor of 0.005199004, given that all other variables are held constant.

For one unit increase in sulphates, the odds of the wine falling at or below quality i multiply by a factor of 15.401872706, given that all other variables are held constant.

For one unit increase in pH, the odds of the wine falling at or below quality i multiply by a factor of 0.228196887, given that all other variables are held constant.

For one unit increase in % alcohol content, the odds of the wine falling at or below quality i multiply by a factor of 2.583998874, given that all other variables are held constant.

### Assumptions
```{r}
# calculate predicted probabilities for each observation
predprobs <- data.frame(predict(final.model,type="probs"))
predprobs %>% filter(row_number() <= 3) # print predictions for first three observations
```

```{r}
# predicted categories
data <- data %>% mutate(pred.comp = predict(final.model,type="class")) # get predicted categories
result <- data %>% group_by(pred.comp,quality_dup) %>% summarise(n=n()) # summarise actual vs. predicted
kable(spread(result,quality_dup,n),format="markdown") # print actual vs. predicted categories
```

### Model Diagnostics
```{r}
data <- data %>% mutate(resid3 = qua3-predprobs[,1],
                        resid4 = qua4-predprobs[,2],
                        resid5 = qua5-predprobs[,3],
                        resid6 = qua6-predprobs[,4],
                        resid7 = qua7-predprobs[,5],
                        resid8 = qua8-predprobs[,6])
data$predor <- predict(final.model, data)
```

```{r}
data %>% count(predor,quality_dup)
confusionMatrix(data$predor,data$quality_dup)
```

```{r}
# binned residual plots of volatile.acidityCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid3,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid4,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid5,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid6,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid7,xlab="volatile.acidityCent")
arm::binnedplot(x=data$volatile.acidityCent,y=data$resid8,xlab="volatile.acidityCent")
```

```{r}
# binned residual plots of citric.acid for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$citric.acid,y=data$resid3,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid4,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid5,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid6,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid7,xlab="citric.acid")
arm::binnedplot(x=data$citric.acid,y=data$resid8,xlab="citric.acid")
```

```{r}
# binned residual plots of chloridesCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$chloridesCent,y=data$resid3,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid4,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid5,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid6,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid7,xlab="chloridesCent")
arm::binnedplot(x=data$chloridesCent,y=data$resid8,xlab="chloridesCent")
```

The binned residual plots show a possible chloridesCent outlier.

```{r}
# binned residual plots of sulphatesCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$sulphatesCent,y=data$resid3,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid4,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid5,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid6,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid7,xlab="sulphatesCent")
arm::binnedplot(x=data$sulphatesCent,y=data$resid8,xlab="sulphatesCent")
```

The binned residual plot shows some potential linear trend and outliers in three categories of quality. We might need to transform sulphatesCent.

```{r}
# binned residual plots of pHCent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$pHCent,y=data$resid3,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid4,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid5,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid6,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid7,xlab="pHCent")
arm::binnedplot(x=data$pHCent,y=data$resid8,xlab="pHCent")
```

There is also some linear trends in the binned residual plots for pHCent. We should also tranform this variable.

```{r}
# binned residual plots of alcohol.Cent for each category of quality
par(mfrow=c(2,3))
arm::binnedplot(x=data$alcoholCent,y=data$resid3,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid4,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid5,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid6,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid7,xlab="alcoholCent")
arm::binnedplot(x=data$alcoholCent,y=data$resid8,xlab="alcoholCent")
```



```{r}
kable(tidy(vif(final.model)))
```


```{r}
# obtain p-values
tidy(coeftest(final.model))
```

p-values of all the variables are extremely small, so they are all significant predictors of the log-odds of the wine falling in or below quality j. We should also not extrapolate beyond quality j=3,4,...,7.

